\documentclass[11pt]{article}

\usepackage[preprint]{neurips_2024}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage{xcolor}

\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black}

\emergencystretch=2em
\setlength{\parskip}{3pt plus 1pt minus 1pt}

\newcommand{\sgmvpi}{$(S, G, E, V, M, \pi)$}
\newcommand{\framework}{\textsc{SciAgent}}

\title{\textbf{Novelty is Socially Constructed:\\
A Multi-Agent Simulation of Scientific Community Dynamics}}

\author{
  Jaesol Shin\\
  \small{WeDataLab}\\
  \texttt{ysys143@wedatalab.com}
}

\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
% TODO: write after experiments are complete
A central open problem in AI Scientist systems is whether novelty evaluation can be automated.
We argue that novelty is not an intrinsic property of a hypothesis but an emergent outcome of
community arbitration — the distributed process by which a scientific community positions a
claim relative to collective knowledge. To test this, we construct a multi-agent simulation
using MiroFish in which heterogeneous reviewer agents with fragmented local memories evaluate
AI-generated hypotheses. We show that (1) isolated semantic classifiers systematically
misidentify novelty when knowledge is distributed across agents, (2) community consensus
accuracy improves non-linearly with knowledge-sharing topology, and (3) human oracle
intervention remains necessary above a noise threshold even with strong local classifiers.
These results provide direct empirical support for the non-automatability claim in
\citet{shin2026sciagent} and constrain the design space of the $V_\text{human}$ layer in
AI Scientist systems.
\end{abstract}

\section{Introduction}
\label{sec:intro}

% Context: the V_human problem
The \framework{} framework \sgmvpi{} \citep{shin2026sciagent} identifies the completeness of
the validator $V$ as the single most consequential design variable in AI Scientist systems.
Among the four validator layers --- $V_\text{syntax}$, $V_\text{semantic}$,
$V_\text{empirical}$, $V_\text{human}$ --- the last is the most resistant to automation.
Prior work has argued informally that novelty evaluation requires human judgment
\citep{lu2024aiscientist}; we provide the first controlled simulation evidence for this claim.

% Gap: no controlled study of V_human
% Approach: MiroFish ABM

\section{Background}
\label{sec:background}

\subsection{Novelty as Community Arbitration}
% V_human is not a single oracle but a distributed process
% Community arbitration: collective positioning of a claim in epistemic space

\subsection{MiroFish: Multi-Agent Social Simulation}
% OASIS-based engine, heterogeneous agents with memory and personality
% Why MiroFish: emergent dynamics, distributed knowledge, partial observability

\section{Simulation Design}
\label{sec:design}

\subsection{Agents}
% Reviewer agents: local M_s, semantic classifier, trust network
% Author agent: submits hypotheses drawn from S
% Editor agent: aggregates reviewer judgments

\subsection{Environment}
% Hypothesis space S: latent concept graph
% Ground truth: global novelty defined over latent classes
% Partial observability: agents see surface form, not latent class

\subsection{Experimental Conditions}
% Condition 1: isolated agents (no memory sharing)
% Condition 2: lattice topology
% Condition 3: small-world network
% Condition 4: fully shared M_s (oracle upper bound)
% Condition 5: human oracle audit (random / high-impact)

\subsection{Metrics}
% False novelty rate, true novelty recall
% Community consensus latency
% Inter-agent agreement (Cohen's kappa)
% Value of human oracle intervention

\section{Results}
\label{sec:results}

% TODO: fill after running simulation

\subsection{Isolated Agents Fail at Global Novelty}

\subsection{Topology Drives Consensus Quality}

\subsection{Human Oracle Threshold}

\section{Discussion}
\label{sec:discussion}

% Implications for V_human design in AI Scientist systems
% Limitations: abstracted hypothesis space, synthetic agents

\section{Conclusion}
\label{sec:conclusion}

\bibliographystyle{abbrvnat}
\bibliography{references}

% ---------------------------------------------------------------
\newpage
\section*{AI Involvement Checklist}

\begin{itemize}
  \item \textbf{Hypothesis development.} The research question (novelty as emergent community
  property, not intrinsic hypothesis property) was proposed by the human author based on
  framework analysis in \citet{shin2026sciagent}. Claude Code helped identify relevant
  Science of Science literature and formalized the agent decision rules.

  \item \textbf{Experimental design.} The human author designed the five experimental
  conditions and the latent concept graph environment. Claude Code implemented the simulation
  codebase (\texttt{src/}) and suggested the inter-agent trust topology variants.

  \item \textbf{Analysis.} Claude Code ran parameter sweeps and generated result figures.
  The human author interpreted results and identified the non-linear threshold behavior.

  \item \textbf{Writing.} Claude Code drafted manuscript sections; the human author wrote
  the core argument and conclusions.
\end{itemize}

% ---------------------------------------------------------------
\newpage
\section*{Responsibility \& Reproducibility Checklist}

\begin{itemize}
  \item \textbf{Claims and evidence.} All quantitative claims are supported by simulation
  results reported in \S\ref{sec:results}. Simulation code is released at
  \texttt{src/} in the supplementary material.

  \item \textbf{Limitations.} The hypothesis space is abstract (graph-structured, not
  drawn from a real scientific domain). Agent personalities are synthetic. Results may
  not transfer directly to real scientific communities.

  \item \textbf{Reproducibility.} Full simulation code, random seeds, and parameter
  configurations are included in supplementary material. All experiments can be
  reproduced on a laptop CPU in under 30 minutes.

  \item \textbf{Data and code availability.} Simulation code released with submission.
  No external datasets used.

  \item \textbf{Ethical and societal impact.} This work informs the design of AI Scientist
  systems. The finding that $V_\text{human}$ cannot be fully automated supports calls for
  maintained human oversight in autonomous research pipelines.
\end{itemize}

\end{document}
